### 2020-03-10 Common/chapters/paper_reproduction.tex
Julien and Johan: paper reproduction proofread

\chapter{Research paper experiment reproduction}
\label{ch:paper_reproduction}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).

Song et al. \cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is one of the biggest dataset available for prostate cancer classification. The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1). The dataset contains a total of 204 patients for the training (labels are known) and 208 patients for the challenge (labels are unknown). This paper explains all the steps to follow with the goal of building a computer-aided diagnosis system for prostate cancer detection. Moreover, it also provides results (AUC) about the algorithm performance on a test set they built themselves (which is not the same as the challenge test set). In fact, they split the PROSTATEx training set into a training set, a validation set and the mentioned test set. These results can be used as a benchmark to compare with the results of the reproduction of the experiment. However, the article gives no information about the results their model achieved on the official PROSTATEx challenge test set (208 patients without labels). Therefore, the reproduction of the experiment will fill that gap by taking part in actual challenge.

This chapter is built in a top-down fashion. First, an overview of the entire process is established in order to understand the purpose of the experiment as a whole. Then, each theoretical notion is described in detail. This includes the structure of the dataset, the steps involved in its processing, and the techniques used as a verification of the proper functioning of the algorithm.Then, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment. This part is followed by the presentation of the raw results the model achieved on the test set (the one built for the experiment), which is itself followed by their analysis in the "Discussion" section. Finally, the last section is devoted to the so-called SPIE-AAPM-NCI Prostate MR Classification Challenge. It presents the results that the model achieved on the challenge test set using the whole training set for the training. The latter section is not part of the original paper.
%and represents the contribution of this work to it.

\section{Process overview}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}. The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.\\
\begin{itemize}

\item The former makes use of the PROSTATEx training set. This dataset is only composed of samples whose clinical significance is provided (labeled data). The first step of the reproduction is the data processing. In fact, many processing steps are involved to transform the original DICOM files into NumPy arrays which can be fed to the neural network. Data processing includes lesion cropping, normalization, MRI images stacking, data augmentation, etc. (see Section \ref{sec:prostatex_data_processing} for further details).\\
Once the processing part is completed, the labeled data are split into a training set, a validation set and a test set using respectively 80\%, 10\% and 10\% of the available data for each subset. The training set samples are then fed to the neural network in batches. 
%, which updates its weight consequently. 
At the end of each epoch, the current model is tested on the validation set and the resulting metrics are plotted on Tensorboard (see \ref{paper_tensorboard}). The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes. Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and is saved. Furthermore, if the validation AUC is decreasing for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a  progressive decrease in validation performance. This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC.\\
As soon as the training phase reaches the defined number of epochs, the model is tested on the test set with the authors' method called "enhanced prediction" to compute the AUC. The latter method consists in predicting an output for each augmented image before averaging the predictions for each lesion of each patient.

\item The second part of the experiment relies on the results of the first one. Indeed, the same data processing pipeline is used and the hyperparameters used to train the neural network are the same. The difference lies in the fact that the entire dataset (i.e. the entire PROSTATEx training set) is used as training set instead of splitting the data into a training/validation/test set. In this way, the model sees more training examples, which maximizes the performance of the model on the challenge test set by increasing its generalization ability. During the training phase, the model is saved at the end of each epoch. Once the training is over, multiple models are chosen around the epoch that produced the best result on the validation set during the first part of the experiment. These models are then evaluated on the challenge test set by submitting their predictions for each lesion to the Grand Challenge organization. The model with the best result on the challenge is considered as the most efficient one.
\end{itemize}
\clearpage

@@

\subsection{PROSTATEx: Data processing}
\label{sec:prostatex_data_processing}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This section describes all processing operations performed on the raw data until it passes through the neural network. All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).

\subsubsection{Dataset description}
\label{sec:prostatex_dataset_description}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website~\cite{33, 34, 35}. This dataset is composed of multiparametric MRIs for a total of 204 training patients and 208 challenge patients. Images were taken under the sagittal, transverse and coronal planes. These MRIs are multiparametric because the same exam was performed using different radiological settings called "sequences". Each sequence allows to see human tissues in a different way on the resulting images. The sequences available within the scope of PROSTATEx are T2 weighted images (T2W), Diffusion weighted images (DWI), Apparent diffusion coefficient (ADC), Dynamic contrast enhanced (DCE), PD (Proton Density) and K-trans. The PROSTATEx dataset comes with two CSV files for the training set:
\begin{itemize}
\item The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance. Multiple findings can be associated with the same patient (ProxID).

@@

\noindent Regarding the challenge patients, two analog CSV files are provided: \textit{ProstateX-Findings-Test.csv} and \textit{ProstateX-Images-Train.csv}. The only difference is the absence of clinical significance. Finally, images are in the DICOM file format (see Section \ref{sec:DICOM}).

@@

\subsubsection{From DICOM to NumPy arrays}
\label{sec:DICOMtoNumPy}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside. Furthermore, DWI images regroup various parameters which create different subcategories. One of them is the so-called "b-value" which is "a factor that reflects the strength and timing of the gradients used to generate diffusion-weighted images. The higher the b-value, the stronger the diffusion effects" \cite{49}. The authors tested three different b-value configurations. This led to the conclusion that using DWI images with the highest b-value increased the performance of the model. For that matter, DWI images with the highest b-value only were used.

Moreover, only images showing the prostate under the transverse plane were used. The reason for this is that the tumors are much more visible under this perspective. 
The first step consisted in converting DICOM files. Algorithm \ref{alg:PROSTATEx_preprocessing} describes the steps involved in converting PROSTATEx's DICOM files to NumPy arrays. The right slices were found thanks to the two CSV files. Important information such as the patient ID, the sequence, the lesion location and the voxel spacing were used as file output names, which allows to use these files independently for the next steps (i.e. without relying on the CSV files). These information are then used for the next step that consists in stacking and augmenting the images. Furthermore, they are also extremely useful for the visualization scripts in order to provide concrete details regarding the images.

@@

An important part of this process is the normalization. In fact, images were normalized before beeing stacked, based on a Z-score, i.e. 
\begin{equation}
\label{eq:normalization}
	Pixel_{{i,j}_{normalized}} = \frac{Pixel_{i,j} - \mu}{\sigma}
\end{equation}
\noindent where $Pixel_{i,j} \in [0,255]$ is a grayscale pixel value, $\mu$ is the mean value of all images of the corresponding sequence for this patient and $\sigma$ the standard deviation of all images of the corresponding sequence for this patient. According to the authors, normalizing each sequence of each patient separately allows to keep slight contrast nuances which ultimately helps the final diagnosis. In other words, all T2-weighted images belonging to a specific patient were concatenated into a single NumPy array. Then, the mean and standard deviation of the array were computed. These values served as $\mu$ and $\sigma$ to normalize the T2-images of this specific patient. Same for the patient's DWI and ADC images. The same process was applied to all patients separately.
Then, the dataset was augmented in the same way as the authors. Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and slid horizontally (value $\in [-1, 0, 1]$ pixel). These techniques alone allowed to create a large enough amount of data. Therefore, two other augmentation methods used by the authors (horizontal stretching by a factor $\in [0.9,1.1]$ and vertical sliding) were not used.

@@

Finally, the data was split into three subsets (training ($80$\%), validation ($10$\%) and test ($10$\%)) and was exported as NumPy arrays. The percentages represent a percentage of patients. Figure \ref{fig:paper_reproduction_split} explains the entire splitting process and gives details about the patients that are in each set. In fact, each subset contains respectively 80\%, 10\% and 10\% of all patients belonging to the "true" and "false" class. At the beginning of the process, class "false" contains $\lfloor 0.8 * 164 \rfloor = 131$ training patients, $\lfloor 0.1 * 164 \rfloor = 16$ test patients and $164 - 131 - 16 = 17$ validation patients. On the other hand, class "true" contains $\lfloor 0.8 * 69 \rfloor = 55$ training patients, $\lfloor 0.1 * 69 \rfloor = 6$ test patients and $69 - 55 - 6 = 8$ validation patients. Then, during the stacking process, some patients were not used since the three sequences were not available for them. These patients were simply discarded. At the end, the "false" class was composed of 131 training patients, 17 validation patients and 15 test patients, whereas the "true" class was composed of 55 training patients, 8 validation patients and 4 test patients. This approach does not take the number of images into account (a patient may have multiple lesions for example) but guarantees that every lesion is exclusive to a specific subset. In other words, it is impossible to find a slighty different version of the same lesion (resulting from the data augmentation) in the training set and in the validation set for example. This would be problematic since the validation set, as well as the test set, need to contain unseen data to be able to evaluate the generalization ability of the model in an unbiased way.

@@

\subsection{Data processing verification}
\label{sec:dataProcessingVerification}
\subsubsection{Cropping verification using red dots}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated. First of all, a red dot was placed on the full image thanks to the "i,j,k" attribute in the CSV file. This full image was exported. Then, images were augmented in the exact same manner as described in section \ref{sec:numpyToAugmentedStacked}. The expected result satisfies the following properties:
\begin{itemize}
	\item The red dot is localized in the center of the cropped and augmented image.
	\item The three cropped and augmented images within a stacked image contain the exact same tissue. 
	\item The three cropped and augmented images within a stacked image are not rotated/shifted/flipped differently. 
	\item A lesion is visible to the naked eye. 
\end{itemize}

\noindent Figure \ref{fig:reddot} shows an example generated during this test. While beeing exactly the same for the 3 images, the red dot looks bigger on the images whose resolution was smaller since it was drawn before resampling the images. These images come from the same exam performed on the same patient on the same day.

@@

\subsubsection{Alignment}
In addition to the previous test, an additional one aiming at checking the alignment of the images was created. This test ensures that the three images show the exact same content at the same spot over the three channels (see Figure \ref{fig:alignment}).

@@

\subsubsection{Architecture}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG). This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 millions images belonging to 1000 classes. Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.

It is first composed of three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks. Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit (ELU) activation function. The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit. The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced by a softmax function for classification.

@@

\subsubsection{Script options}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}
Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary. This requirement was solved thanks to the Python module called "argparse". The latter automatically generates help/usage messages and displays errors when the arguments typed in by the user are invalid. Table \ref{fig:paper_reproduction_options} shows all options accepted by the script, provides information about what they concretely mean and indicates their respective types.

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[!h]
\resizebox{\textwidth}{!}{%
\begin{tabular}{|llll|}
\hline
\rowcolor[HTML]{DAE8FC} 
\textbf{Command}            & \textbf{Description}                                                                                                                                                                                                & \textbf{Required}                                                           & \textbf{Type} \\ \hline
\textbf{--trainingset}     & Training set  path                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
\textbf{--validationset}   & Validation set path                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
\textbf{--batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Int           \\ \hline
\textbf{--nbepochs}         & Number of epochs the training phase has to last                                                                                                                                                                     & True                                                                        & Int           \\ \hline
\textbf{--lr}               & Learning rate used by the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 1e-3\end{tabular}               & Float         \\ \hline
\textbf{--lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
\textbf{--cudadevice}      & GPU name to run the experiment                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
\textbf{--modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ''\end{tabular}                 & String        \\ \hline
\textbf{--dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 0.3\end{tabular}                & Float         \\ \hline

\textbf{--inputchannel}     & \begin{tabular}[c]{@{}l@{}}Number of channels of the input images\\{[}1,3{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}True\end{tabular} & Int        \\ \hline

\textbf{--optimizedmetric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
\textbf{--outputdirectory}  & Root of the output directory used by Tensorboard to save the models                                                                                                                                                 & True                                                                        & String        \\ \hline
\end{tabular}%
}
\caption{Complete list of script options}
\label{fig:paper_reproduction_options}
\end{table}

\subsubsection{Tensorboard}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Tensorboard's official website \cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
\begin{itemize}
\item Tracking and visualizing metrics.
\item Visualizing the model graph (ops and layers).
\item Viewing histograms of weights, biases, or other tensors as they change over time.
\item Projecting embeddings to a lower dimensional space.
\item Displaying images, text, and audio data"\cite{39}.
\end{itemize}
In this experiment, Tensorboard is mainly used to plot Matplotlib figures of the model performance. In fact, during the training the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC. These metrics are computed separately on the training and on the validation sets at the end of each epoch. They are then stored and plotted on the same figure at the end of the process thanks to Tensorboard. In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
\label{paper_tensorboard}

>>>>

\chapter{Research paper experiment reproduction}
\label{ch:paper_reproduction}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This chapter is based on the article "Computer-Aided Diagnosis of Prostate Cancer Using a Deep Convolutional Neural Network from Multiparametric MRI" from Song et al.~\cite{07}, shortly presented in Chapter \ref{ch:literature_review}. It aims at reproducing the experiment of the paper in order to acquire the medical, theoretical and technical background before proposing a transfer learning method as a way to improve the classification using other body parts (see Chapter \ref{ch:transfer_learning}).

Song et al.~\cite{07} proposed a deep convolutional neural network (DCNN) method to detect prostate cancer based on the SPIE-AAPM-NCI PROSTATEx Challenge dataset. This dataset is one of the biggest datasets available for prostate cancer classification. The two different output classes of the latter are benign lesion (class 0) and malignant lesion (class 1). The dataset contains a total of 204 patients for the training (labels are known) and 208 patients for the challenge (labels are unknown). This paper explains all the steps to follow with the goal of building a computer-aided diagnosis system for prostate cancer detection. Moreover, it also provides results (AUC) about the algorithm performance on a test set they built themselves (which is not the same as the challenge test set). In fact, they split the PROSTATEx training set into a training set, a validation set and the mentioned test set. These results can be used as a benchmark to compare with the results of the reproduction of the experiment. However, the article gives no information about the results their model achieved on the official PROSTATEx challenge test set (208 patients without labels). Therefore, the reproduction of the experiment will fill that gap by taking part in the actual challenge.

This chapter is built in a top-down fashion. First, an overview of the entire process is established in order to understand the purpose of the experiment as a whole. Then, each theoretical notion is described in detail. This includes the structure of the dataset, the steps involved in its processing, and the techniques used as a verification of the proper functioning of the algorithm. Then, the training phase is dealt with in depth and all hyperparameters, options and implementation choices are described to ensure the reproductibility of the experiment. This part is followed by the presentation of the raw results the model achieved on the test set (the one built for the experiment), which is itself followed by their analysis in the "Discussion" section. Finally, the last section is devoted to the so-called SPIE-AAPM-NCI Prostate MR Classification Challenge. It presents the results that the model achieved on the challenge test set using the whole training set for the training. The latter section is not part of the original paper.
%and represents the contribution of this work to it.

\section{Process overview}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Schematically speaking, the entire experiment process can be represented as shown on Figure \ref{fig:paper_reproduction_process}. The first part of the experiment is about the reproduction of the paper experiment, whereas the second one is devoted to the participation in the SPIE-AAPM-NCI PRO\- STATEx challenge since the authors of the article did not take part in it.

\subsubsection{Experiment reproduction}
The first part of the experiment makes use of the PROSTATEx training set. This dataset is only composed of samples whose clinical significance is provided (labeled data). The first step of the reproduction is the data processing. In fact, many processing steps are involved to transform the original DICOM files into NumPy arrays which can be fed to the neural network. Data processing includes lesion cropping, normalization, MRI images stacking, data augmentation, etc. (see Section \ref{sec:prostatex_data_processing} for further details).\\
Once the processing part is completed, the labeled data are split into a training set, a validation set and a test set using respectively 80\%, 10\% and 10\% of the available data for each subset. The training set samples are then fed to the neural network in batches. 
%, which updates its weight consequently. 
At the end of each epoch, the current model is tested on the validation set and the resulting metrics are plotted in Tensorboard (see Section \ref{paper_tensorboard}). The validation set is independent of the training set and has never been seen by the model, which gives an indication of how well the latter generalizes. Then, if the current model reached a higher AUC and a higher accuracy than the previous best model, it becomes the new best model and it is saved. Furthermore, if the validation AUC decreases for a defined number of epochs, the learning rate is reduced to prevent overfitting and avoid a progressive decrease in validation performance. This technique contributes to maintaining a good generalization level and optimizes the training with respect to the AUC.\\
As soon as the training phase reaches the defined number of epochs, the model is tested on the test set with the authors' method called "enhanced prediction" to compute the AUC. The latter method consists in predicting an output for each augmented test image before averaging the predictions for each lesion of each patient.

\subsubsection{PROSTATEx challenge}
The second part of the experiment relies on the results of the first one. Indeed, the same data processing pipeline is applied and the hyperparameters used to train the neural network are the same. The difference lies in the fact that the entire dataset (i.e. the entire PROSTATEx training set) is used as training set instead of splitting the data into a training, validation and test sets. In this way, the model sees more training examples, which maximizes the performance of the model on the challenge test set by increasing its generalization ability. During the training phase, the model is saved at the end of each epoch. Once the training is over, multiple models are chosen around the epoch that produced the best result on the validation set during the first part of the experiment. These models are then evaluated on the challenge test set by submitting their predictions for each lesion to the Grand Challenge organization. The model with the best result on the challenge is considered as the most efficient one.

@@

\subsection{PROSTATEx: Data processing}
\label{sec:prostatex_data_processing}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}This section describes the processing operations performed on the raw data until it passes through the neural network. All steps mentioned below are also valid for the second part of the experiment about the challenge (Section \ref{sec:classification_challenge}).

\subsubsection{Dataset description}
\label{sec:prostatex_dataset_description}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}The SPIE-AAPM-NCI PROSTATEx Challenge dataset is publicly available on the Cancer Imaging Archive website~\cite{33, 34, 35}. This dataset is composed of multiparametric MRIs for a total of 204 training patients and 208 challenge patients. Images were taken under the sagittal, transverse and coronal planes. These MRIs are multiparametric because the same exam was performed using different radiological settings called "sequences". Each sequence allows to see human tissue in a different way on the resulting images. The sequences available within the scope of PROSTATEx are T2 weighted images (T2W), Diffusion weighted images (DWI), Apparent diffusion coefficient (ADC), Dynamic contrast enhanced (DCE), PD (Proton Density) and K-trans. The PROSTATEx dataset comes with two CSV files for the training set:
\begin{itemize}
\item The first one, \textit{ProstateX-Findings-Train.csv}, lists all findings with their clinical significance. Multiple findings can belong to the same patient (ProxID).

@@

\noindent Regarding the challenge patients, two analog CSV files are provided: \textit{ProstateX-Findings-Test.csv} and \textit{ProstateX-Images-Train.csv}. The only difference is the absence of clinical significance. Finally, images are in the DICOM file format (see \mbox{Section \ref{sec:DICOM}}).

@@

\subsubsection{From DICOM to NumPy arrays}
\label{sec:DICOMtoNumPy}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}Before anything else, T2W, DWI and ADC grayscale images were used, which means that DCE, PD and Ktrans images were left aside. Furthermore, DWI images regroup various parameters which create different subcategories. One of them is the so-called "b-value" which is "a factor that reflects the strength and timing of the gradients used to generate diffusion-weighted images. The higher the b-value, the stronger the diffusion effects"~\cite{49}. The authors tested three different b-value configurations. This led to the conclusion that using DWI images with the highest b-value increased the performance of the model. For that matter, DWI images with the highest b-value only were used in this work. Moreover, only images showing the prostate under the transverse plane were used. The reason for this is that the tumors are much more visible under this perspective.

The first step consisted in converting PROSTATEx's DICOM files to NumPy arrays. Algorithm \ref{alg:PROSTATEx_preprocessing} describes the steps involved in this process. The right slices were found thanks to the two CSV files. Important information such as the patient ID, the sequence, the lesion location and the voxel spacing was included in the file output names, which allows to use these files independently for the next steps (i.e. without relying on the CSV files) that consist in stacking and augmenting the images. Furthermore, this information is also extremely useful for the visualization and verification scripts in order to provide concrete details regarding the displayed images.

@@

An important part of this process is the normalization. In fact, images were normalized before beeing stacked, based on a Z-score, i.e. 
\begin{equation}
\label{eq:normalization}
	Pixel_{{i,j}_{normalized}} = \frac{Pixel_{i,j} - \mu}{\sigma}
\end{equation}
\noindent where $Pixel_{i,j} \in [0,255]$ is a grayscale pixel value, $\mu$ is the mean value of all images of the corresponding sequence for this patient and $\sigma$ the standard deviation of all images of the corresponding sequence for this patient. According to the authors, normalizing each sequence of each patient separately allows to keep slight contrast nuances which ultimately help the final diagnosis. In other words, all T2-weighted images belonging to a specific patient were concatenated into a single NumPy array. Then, the mean and standard deviation of the array were computed. These values served as $\mu$ and $\sigma$ to normalize the T2-images of this specific patient. The equivalent was applied to the patient's DWI and ADC images, and the same process was employed for all patients separately.
Then, the dataset was augmented in the same way as the authors. Concretely, images were rotated ($-20$ to $20^\circ$), flipped horizontally (probability of $0.5$) and shifted horizontally (value $\in [-1, 0, 1]$ pixel). These techniques alone allowed to create a large enough amount of data. Therefore, two other augmentation methods used by the authors (horizontal stretching by a \mbox{factor $\in [0.9,1.1]$} and vertical shifting) were not used.

@@

Finally, the data was split into three subsets (training ($80$\%), validation ($10$\%) and test ($10$\%)) and was exported as NumPy arrays. The percentages represent a percentage of patients. Figure \ref{fig:paper_reproduction_split} explains the entire splitting process and gives details about the patients that are in each set. In fact, each subset contains respectively 80\%, 10\% and 10\% of all patients belonging to the "true" and "false" class. At the beginning of the process, class "false" contains $\lfloor 0.8 * 164 \rfloor = 131$ training patients, $\lfloor 0.1 * 164 \rfloor = 16$ test patients and $164 - 131 - 16 = 17$ validation patients. On the other hand, class "true" contains $\lfloor 0.8 * 69 \rfloor = 55$ training patients, $\lfloor 0.1 * 69 \rfloor = 6$ test patients and $69 - 55 - 6 = 8$ validation patients. Then, during the stacking process, some patients were not used since the three sequences were not available for them. These patients were simply discarded. At the end, the "false" class contained 131 training patients, 17 validation patients and 15 test patients, whereas the "true" class contained 55 training patients, 8 validation patients and 4 test patients. This approach does not take the number of images into account (a patient may have multiple lesions for example) but guarantees that every lesion is exclusive to a specific subset. In other words, it is impossible to find a slighty different version of the same lesion (resulting from the data augmentation) in the training set and in the validation set for example. This would be problematic since the validation set, as well as the test set, need to contain unseen data to be able to evaluate the generalization ability of the model in an unbiased way.

@@

\subsection{Data processing verification}
\label{sec:dataProcessingVerification}
\subsubsection{Cropping verification using red dots}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}In order to check the correctness of the augmentation process, images with a red dot at the lesion coordinates were generated. First of all, a red dot was placed on the full image thanks to the "i,j,k" attribute of the CSV file. This full image was exported. Then, images were augmented in the exact same manner as described in Section \ref{sec:numpyToAugmentedStacked}. The expected result satisfies the following properties:
\begin{itemize}
	\item The red dot is localized in the center of the cropped and augmented image.
	\item The three cropped and augmented images within a stacked image contain the exact same tissue. 
	\item The three cropped and augmented images within a stacked image are not rotated/shifted/flipped differently. 
	\item A lesion is visible to the naked eye. 
\end{itemize}

\noindent Figure \ref{fig:reddot} shows an example generated during this test. While being exactly the same size for the 3 images, the red dot looks bigger on the images whose resolution was smaller since it was drawn before resampling the images. These images come from the same exam performed on the same patient on the same day.

@@

\subsubsection{Alignment}
In addition to the previous test, another one aiming at checking the alignment of the images was implemented. This test ensures that the three images show the exact same content at the same spot over the three channels (see Figure \ref{fig:alignment}).

@@

\subsubsection{Architecture}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}The model architecture used in the paper is a modified version of the VGG network from the Oxford's Visual Geometry Group (VGG). This model was initially designed as part of the Large Scale Visual Recognition Challenge 2014 that used the ImageNet dataset composed of 14 million images belonging to 1000 classes. Figure \ref{fig:paper_model} illustrates its structure and its corresponding implementation in Python with the PyTorch framework.

It is first composed of three convolution-dropout-max-pooling blocks followed by three fully-connected-dropout blocks. Each convolutional box (in blue) in the figure represents in reality three layers: the convolutional layer, the batch normalization layer and the exponential linear unit (ELU) activation function. The same principle applies to the fully connected layer box (in orange) that is divided into a fully connected layer followed by the exponential linear unit. The last fully connected box (in purple) has the same structure except that the exponential linear unit is replaced with a softmax function for classification.

@@

\subsubsection{Script options}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Since training a neural network requires datasets, hyperparameters and many more configuration choices, the creation of a generic script that accepts multiple options is necessary. This requirement was solved thanks to the Python module called "argparse". The latter automatically generates help/usage messages and displays errors when the arguments typed in by the user are invalid. Table \ref{fig:paper_reproduction_options} shows all options accepted by the script, provides information about what they concretely mean and indicates their respective types.

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[!h]
\resizebox{\textwidth}{!}{%
\begin{tabular}{|llll|}
\hline
\rowcolor[HTML]{DAE8FC} 
\textbf{Command}            & \textbf{Description}                                                                                                                                                                                                & \textbf{Required}                                                           & \textbf{Type} \\ \hline
\textbf{-{}-trainingset}     & Training set  path                                                                                                                                                                                                  & True                                                                        & String        \\ \hline
\textbf{-{}-validationset}   & Validation set path                                                                                                                                                                                                 & True                                                                        & String        \\ \hline
\textbf{-{}-batchsize}        & Number of samples per batch                                                                                                                                                                                         & True                                                                        & Int           \\ \hline
\textbf{-{}-nbepochs}         & Number of epochs the training phase has to last                                                                                                                                                                     & True                                                                        & Int           \\ \hline
\textbf{-{}-lr}               & Learning rate used by the optimizer                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 1e-3\end{tabular}               & Float         \\ \hline
\textbf{-{}-lossfunction}     & \begin{tabular}[c]{@{}l@{}}Loss function name\\{[}CrossEntropyLoss, L1Loss, MSELoss{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'CrossEntropyLoss'\end{tabular} & String        \\ \hline
\textbf{-{}-cudadevice}      & GPU name to run the experiment                                                                                                                                                                                      & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'cuda'\end{tabular}             & String        \\ \hline
\textbf{-{}-modeltoload}      & \begin{tabular}[c]{@{}l@{}}Pretrained model name \\If given, load it, otherwise randomly initialize it\end{tabular}                                                                                     & \begin{tabular}[c]{@{}l@{}}False\\ Default: ''\end{tabular}                 & String        \\ \hline
\textbf{-{}-dropout}          & Dropout probability                                                                                                                                                                                                 & \begin{tabular}[c]{@{}l@{}}False\\ Default: 0.3\end{tabular}                & Float         \\ \hline

\textbf{-{}-inputchannel}     & \begin{tabular}[c]{@{}l@{}}Number of channels of the input images\\{[}1,3{]}\end{tabular}                                                                                                            & \begin{tabular}[c]{@{}l@{}}True\end{tabular} & Int        \\ \hline

\textbf{-{}-optimizedmetric} & \begin{tabular}[c]{@{}l@{}}Metric to optimize\\ The best model during the training will be saved according to it\\{[}'auc', 'accuracy', 'precision', 'recall', 'f1score', 'specificity'{]}\end{tabular} & \begin{tabular}[c]{@{}l@{}}False\\ Default: 'auc'\end{tabular}              & String        \\ \hline
\textbf{-{}-outputdirectory}  & Root of the output directory used by Tensorboard to save the models                                                                                                                                                 & True                                                                        & String        \\ \hline
\end{tabular}%
}
\caption{Complete list of script options}
\label{fig:paper_reproduction_options}
\end{table}

\subsubsection{Tensorboard}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}Tensorboard's official website~\cite{39} claims that "Tensorboard provides the visualization and tooling needed for machine learning experimentation:
\begin{itemize}
\item Tracking and visualizing metrics.
\item Visualizing the model graph (ops and layers).
\item Viewing histograms of weights, biases, or other tensors as they change over time.
\item Projecting embeddings to a lower dimensional space.
\item Displaying images, text, and audio data"~\cite{39}.
\end{itemize}
In this experiment, Tensorboard is mainly used to plot Matplotlib figures of the model performance. In fact, during the training the following metrics are computed: loss, accuracy, precision, recall, F1-score, specificity and AUC. These metrics are computed separately on the training and on the validation sets at the end of each epoch. They are then stored and plotted on the same figure at the end of the process thanks to Tensorboard. In addition to this, written reports regarding which model was the best and the results it achieved are also added to the dashboard.
\label{paper_tensorboard}

\subsubsection{Model roulette}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Cl{\'e}ment}}As stated by the authors, "the training process is sensitive to the parameter initialization" \cite{07}. Experience showed us that the exact same hyperparameters with two different initializations could give opposite results. Therefore, in order not to train the model vainly, a script which generates a given number of initializations was created. Each untrained model was then tested on the validation set since the best model during training is saved according to its performance on the validation set. The model whose score was the highest on a given metric (the AUC in this experiment) was saved and used as base model for the experiments. It can then be used as initialization model using the script option \textit{-{}-modeltoload <path>}.
